---
title: "report"
author: "An Vo"
date: "Ngày 14 tháng 03 năm 2017"
output: html_document
---

## Introduction  
This project uses data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which the participants did the barbell lifts exercise. This document describes the predictive model for the Weight Lifting Exercises Dataset. The goal is to fit a predictive model to the provided data in order to make prediction.

## Libraries loading
```{r, cache = T}
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(corrplot)
```
### Data download
```{r, cache = T}
trainUrl <-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
trainFile <- "./data/pml-training.csv"
testFile  <- "./data/pml-testing.csv"
if (!file.exists("./data")) {
  dir.create("./data")
}
if (!file.exists(trainFile)) {
  download.file(trainUrl, destfile=trainFile, method="curl")
}
if (!file.exists(testFile)) {
  download.file(testUrl, destfile=testFile, method="curl")
}
```  
### Read the Data
Read the downloaded csv files into two data frames.  
```{r, cache = T}
trainRaw <- read.csv("./data/pml-training.csv")
testRaw <- read.csv("./data/pml-testing.csv")
dim(trainRaw)
dim(testRaw)
```
The training set contains 19622 observations and 160 variables, while the test set contains 20 observations and 160 variables. The "classe" variable in the training set is the outcome to predict. 

### Data cleansing
We will clean the data to get rid of observations with missing values as well as some meaningless variables.
```{r, cache = T}
sum(complete.cases(trainRaw))
```
Remove columns that contain NA missing values.
```{r, cache = T}
trainRaw <- trainRaw[, colSums(is.na(trainRaw)) == 0] 
testRaw <- testRaw[, colSums(is.na(testRaw)) == 0] 
```  
Get rid of some columns that do not contribute much to the accelerometer measurements.
```{r, cache = T}
classe <- trainRaw$classe
trainRemove <- grepl("^X|timestamp|window", names(trainRaw))
trainRaw <- trainRaw[, !trainRemove]
trainCleaned <- trainRaw[, sapply(trainRaw, is.numeric)]
trainCleaned$classe <- classe
testRemove <- grepl("^X|timestamp|window", names(testRaw))
testRaw <- testRaw[, !testRemove]
testCleaned <- testRaw[, sapply(testRaw, is.numeric)]
```
The cleaned training data set contains 19622 observations and 53 variables. The testing data set contains 20 observations and 53 variables. The "classe" variable is still in the cleaned training set.

### Slicing the data
We split the cleaned training set into a pure training data set (70%) and a validation data set (30%). We will use the validation data set to perform cross validation in future steps.  
```{r, cache = T}
set.seed(22519) # For reproducibile purpose
inTrain <- createDataPartition(trainCleaned$classe, p=0.70, list=F)
trainData <- trainCleaned[inTrain, ]
testData <- trainCleaned[-inTrain, ]
```

## Data Modeling
We fit a predictive model for activity recognition using Random Forest algorithm because it automatically selects important variables and is robust to correlated covariates & outliers in general. We will use 5-fold cross validation when applying the algorithm.  
```{r, cache = T}
controlRf <- trainControl(method="cv", 5)
modelRf <- train(classe ~ ., data=trainData, method="rf", trControl=controlRf, ntree=250)
modelRf
```
Then, we estimate the performance of the model on the validation data set.  
```{r, cache = T}
predictRf <- predict(modelRf, testData)
confusionMatrix(testData$classe, predictRf)
```
```{r, cache = T}
accuracy <- postResample(predictRf, testData$classe)
accuracy
outOfSampleError <- 1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1])
outOfSampleError
```

## Prediction on Test Data Set
Now, we apply the model to the original testing data set downloaded from the data source. We remove the `problem_id` column first.  
```{r, cache = T}
result <- predict(modelRf, testCleaned[, -length(names(testCleaned))])
result
```